---
title: "ECON 452 Final Project"
author: "Kevin Tan, Rachel Schonbaum, Jessica Lim, Shubhangi Kumari"
date: "November 20, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreach)
library(tidyverse)
library(glmnet)
library(caret)
library(randomForest)
library(magrittr)
library(do)
library(readxl)
library(writexl)
library(arrow)
library(stringr)
library(pROC)
library(ranger)
library(gridExtra)
library(plotly)
library(akima)
library(reshape2)
library(SmartEDA)
library(psych)
```


```{r}
load('test.Rdata')
test = test %>% as.data.frame()
sum(test$part_athletics == 0)
test
```

```{r}
testDF = test %>% mutate(score = math_test + reading_test + science_test,
                         region = factor(region, levels = c(1, 2, 3, 4), 
                                labels = c('northeast', 'midwest', 'south', 'west')), 
                         race = factor(race, levels = 1:4,
                                       labels = c('white','black','hispanic', 'other')),
                gender = factor(gender, levels = c(1, 2), 
                                labels = c('male', 'female')),
                mom_educ = factor(mom_educ, levels = 1:5, 
                                  labels = c('1to8', '9to11', 'hs', 'some_college', 'college_grad')),
                home_language_nonenglish = factor(home_language_nonenglish, levels=c(0, 1),
                                                  labels = c('english', 'nonenglish')),
                both_parents = factor(both_parents, levels=c(0, 1),
                                      labels = c('not_both_parents', 'both_parents')),
                mom_curr_married = factor(mom_curr_married, levels=c(0, 1),
                                          labels = c('mom_not_married', 'mom_married')),
                mom_married_at_birth = factor(mom_married_at_birth, levels = c(0, 1),
                                              labels = c('mom_not_married_at_birth', 'mom_married_at_birth')),
                family_type = factor(family_type, levels = 1:5,
                                     labels = c('2_parents_and_siblings', '2_parents_no_siblings', 
                                                '1_parent_and_siblings', '1_parent_no_siblings',
                                                'other_family_type')),
                part_dance = factor(part_dance, levels = c(0, 1),
                                    labels = c('no_dance', 'dance')),
                part_art = factor(part_art, levels = c(0, 1),
                                    labels = c('no_art', 'art')),
                part_music = factor(part_music, levels = c(0, 1),
                                    labels = c('no_music', 'music')),
                part_club = factor(part_club, levels = c(0, 1),
                                   labels = c('no_club', 'club')),
                part_athletics = factor(part_athletics, levels = c(0, 1),
                                        labels = c('no_athletics', 'athletics')),
                mom_work_status = factor(mom_work_status, levels = 1:3,
                                         labels = c('mom_works_35_or_more', 
                                                    'mom_works_less_than_35', 'mom_no_work')),
                has_library_card = factor(has_library_card, levels = c(0, 1),
                                          labels = c('no_library_card', 'library_card')),
                has_home_computer = factor(has_home_computer, levels = c(0, 1),
                                           labels = c('no_home_computer', 'home_computer')),
                problem_crowding = factor(problem_crowding, levels=c(0,1),
                                          labels=c('no_overcrowding', 'overcrowding')),
                problem_turnover = factor(problem_turnover, levels=c(0,1),
                                          labels=c('no_turnover', 'turnover')),
                problem_gangs = factor(problem_gangs, levels=c(0, 1),
                                       labels = c('no_gangs', 'gangs')),
                problem_crime = factor(problem_crime, levels=c(0, 1),
                                       labels = c('no_crime', 'crime')),
                problem_drugs = factor(problem_drugs, levels = c(0, 1),
                                       labels = c('no_drugs', 'drugs')),
                problem_attacks = factor(problem_attacks, levels = c(0, 1),
                                        labels = c('no_attack', 'attack')),
                problem_parents = factor(problem_parents, levels = c(0, 1),
                                         labels = c('no_parents_prob', 'parents_prob')),
                problem_weapons = factor(problem_weapons, levels=c(0,1),
                                         labels=c('no_weapons', 'weapons')),
                pct_minority = factor(pct_minority, levels = 1:5,
                                      labels = c('less_than_10_minority', '10_to_25_minority',
                                                 '25_to_50_minority', '50_to_75_minority',
                                                 '75_or_more_minority')),
                school_type = factor(school_type, levels = 1:3,
                                     labels = c('public', 'catholic', 'other_private')),
                school_has_security = factor(school_has_security, levels = c(0, 1),
                                             labels = c('no_security', 'security'))
                ) %>% as.data.frame()
ExpCustomStat(testDF, Cvar=names(testDF)[sapply(testDF, is.factor)], 
              gpby = FALSE)
#%>% write_xlsx('summaryStatisticsCategorical.xlsx')
```

```{r}
testDF = data.frame(testDF)
num_cols = names(testDF)[sapply(testDF, is.numeric)]
sumStats = testDF[,num_cols] %>% describe
#sumStats %>% mutate(names = rownames(sumStats))
#sumStats %>% write_xlsx('summaryStatistics.xlsx')
sumStats

```


```{r}
testDF
```




Test 1 and 3

```{r}
testDF %>% mutate(part_extrac = )
partDF = testDF %>% 
  mutate(no_part_dance = part_dance == 'no_dance',
         no_part_athletics = part_athletics == 'no_athletics', 
         no_part_club = part_club == 'no_club',
         no_part_music = part_music == 'no_music',
         no_part_art = part_art == 'no_art') %>%
  mutate(some_part_extra = !(no_part_dance & no_part_athletics & no_part_club & no_part_music & no_part_art)) %>% 
  select(-no_part_dance, -no_part_athletics, -no_part_club, -no_part_music, -no_part_art)  %>% 
  mutate(part_total = (part_dance=='dance') + (part_athletics=='athletics') + 
                    (part_club=='club') + (part_music=='music') + (part_art=='art'))
lm_reduced = lm(score ~ part_total + 
           family_income + mom_work_status + both_parents + mom_educ + problem_crime, data = partDF)
lm_reduced %>% summary
```

The t-stat on some_part_extraTRUE tests $H_0 : partExtra = 0$, or whether there is any effect of extracurriculars on test scores at all.


```{r}
lm_full = lm(score ~ part_dance + part_athletics + part_club + part_music + part_art + 
           family_income + mom_work_status + both_parents + mom_educ + problem_crime, data = partDF)
lm_full %>% summary(correlation=TRUE, symbolic.cor=TRUE)

```
```{r}
partDF %>% select(score) %>% max
```

Now, we want to test if it matters at all what these extracurriculars are. Test H0, all CCAs don't matter, vs H1, there exists an extracurricular 

```{r}
sum(lm_full$residuals ^ 2) # 14 params 1 intercept
sum(lm_reduced$residuals ^ 2) # 10 params 1 intercept
((sum(lm_reduced$residuals ^ 2) - sum(lm_full$residuals ^ 2)) / 4) / (sum(lm_full$residuals ^ 2) / 8090)
pf(21.64233, 4, 8090) # R rounds output to 1
qf(0.95, 4, 8090) # need 2.3703 for statsig at 5% level
```

$F = \frac{(SSR_r - SSR_u) / q}{SSR_u /(n-k-1)} = \frac{(SST(1-R^2_r) - SST(1-R^2_u)) / q}{SST(1-R^2_u)/(n-k-1)} = \frac{1-R^2_r - 1 + R^2_u}{q(1-R^2_u)/(n-k-1)} = \frac{(R^2_u - R^2_r)/q}{(1-R^2_u)/(n-k-1)}$, as $SSR = SST-SSE = SST-R^2(SST) = SST(1-R^2)$

```{r}
lm_controls = lm(score ~ family_income + mom_work_status + both_parents + mom_educ + problem_crime, data = partDF)
lm_controls %>% summary
```

```{r}
lm_multiple = lm(score ~ part_one + part_multiple +
           family_income + mom_work_status + both_parents + mom_educ + problem_crime, 
   data = partDF %>% mutate(part_one = as.numeric(part_total == 1), part_multiple = as.numeric(part_total > 1))) 
lm_multiple %>% summary

```


```{r}
lm_extra = lm(score ~ part_dance + part_athletics + part_club + part_music + part_art + 
           family_income + mom_work_status + both_parents + mom_educ + problem_crime, data = partDF)
lm_extra %>% summary
```

```{r}
sum(lm_extra$residuals ^ 2) # 14 params 1 intercept
sum(lm_controls$residuals ^ 2) # 10 params 1 intercept
((sum(lm_controls$residuals ^ 2) - sum(lm_extra$residuals ^ 2)) / 5) / (sum(lm_extra$residuals ^ 2) / 8090) # 41.41747
pf( 41.41747, 5, 8090) # R rounds output to 1
qf(0.95, 5, 8090) # need  2.37303 for statsig at 5% level
```

```{r}
stargazer(lm_reduced, lm_full, type='text')
```


```{r}
stargazer(lm_controls, lm_full, type='text')
```


```{r}
stargazer(lm_multiple, type='text')

```


```{r}
plot(lm_full)
plot(lm_controls)
plot(lm_extra)
```



```{r}
lm(score ~ part_dance + part_athletics + part_club + part_music + part_art + 
           family_income + mom_work_status + both_parents + mom_educ + problem_crime, data = partDF) %>% summary()
```


```{r}
partDF %>% select(some_part_extra, part_dance, part_athletics, part_club, part_music, part_art) %>% mutate(part_dance = part_dance=='dance', part_athletics = part_athletics=='athletics', part_club = part_club=='club', part_music = part_music=='music', part_art = part_art=='art') %>% rankMatrix
```

$SomePartEC := 1-\prod_i^q (1-PartEC_i)$

```{r}
data = testDF %>% select(-reading_test, -math_test, -science_test)
trainID = sample(nrow(data), size = floor(0.7*nrow(data)))
trainSet = data[trainID,]
testSet = setdiff(data, trainSet)

trainMat = model.matrix(score ~ ., trainSet)[,-1]
testMat = model.matrix(score ~ ., testSet)[,-1]

cvNet = cv.glmnet(trainMat, trainSet$score, alpha=0.99, 
                      family="gaussian")
lambda = cvNet$lambda.min
lassoReg = glmnet(trainMat, trainSet$score, alpha=0.99, family="gaussian", 
                  lambda = 1)
plot(cvNet)
lambda
lassoReg$beta

# family_income + mom_work_status + both_parents + mom_educ + problem_crime

# gender, race, mom_educ, siblings, pct_minority, extracurriculars, 
# home_language_nonenglish, problem_drugs, problem_gangs, problem_attack,
# has_library_card, has_home_computer, mom_curr_married
```

```{r}
unifs = runif(1000)
hist(-2*unifs/(1+(2^2)*unifs)+(1/2^2) * log(1+(2^2)*unifs))
```

```{r}
hist(-log(1+unifs))
```
```{r}

```








Indep is composite score
Dep is all extracurriculars, 



trainMat = model.matrix(Shariah ~ ., trainSet)[,-1]
  testMat = model.matrix(Shariah ~ ., testSet)[,-1]
  sumNeg = length(trainSet$Shariah[trainSet$Shariah==0])
  sumPos = length(trainSet$Shariah) - sumNeg
  weights = numeric(length(trainSet$Shariah))
  weights[trainSet$Shariah==0] = rep(1-sumNeg/(sumNeg+sumPos), sum(trainSet$Shariah==0))
  weights[trainSet$Shariah==1] = rep(1-sumPos/(sumNeg+sumPos), sum(trainSet$Shariah==1))
  #c(1-sumNeg/(sumNeg+sumPos), 1-sumPos/(sumNeg+sumPos))
  if(!lasso){
    lambda=0
  }
  else{
    cvNet = cv.glmnet(trainMat, as.factor(trainSet$Shariah), alpha=1, 
                      family="binomial", weights=weights)
    lambda = cvNet$lambda.min
  }
  lassoLog = glmnet(trainMat, as.factor(trainSet$Shariah), alpha=1, family="binomial", 
                  lambda = lambda, weights=weights)
  lassoLogTestProb = predict(lassoLog, testMat, type="response")




